Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural music skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-14 abstract/abstract.tex
Updates from Overleaf

The abstract should have at least 200 but not more than 600 words. Placed on the right page next to a blank left page.

\bigskip
Keywords: Music; Music Theory; Music Information Retrieval; Music Structure Analysis; Machine Learning; Neural Networks; Deep Learning

>>>>

Music, an integral part of human culture, offers a diverse yet complex field of study due to its intricate styles, mutable ground truth, and subjective nature. This research, conducted in collaboration with Epidemic Sound AB and the Music Technology Group (MTG) at Universitat Pompeu Fabra (UPF), employs self-supervised contrastive learning of musical representation to uncover the fundamental structure of Western tonal music.

Embracing a subtle novel approach, this process is facilitated by applying Siamese networks with a triplet loss, mimicking human aural skills in interpreting music to discern abstract and semantic musical elements, irrespective of the sonic qualities. The study replaces traditional acoustical features with deep audio embeddings to compute high-level, sound-agnostic, and content-sensitive music identification.

Results XXXXXX

The preliminary results suggest that our approach to learning music-informed embeddings holds significant potential for nearly all MIR downstream tasks. Music-motivated embeddings represent a promising technique, adaptable potentially to other tasks hindered by data scarcity. This method could significantly influence advancements in intelligent music recommendation systems and the efficient enforcement of intellectual property rights.

\bigskip
Keywords: Music Information Retrieval; Music Structure Analysis; Deep Audio Embeddings, Aural Skills

### 2023-05-14 evaluation/subsections/beyond dsp.tex
Updates from Overleaf

>>>>

\section{\textit{Embeddiogram} custom feature}

Features extracted from audio files play a vital role in characterizing and differentiating sound and music. One such feature, known as the \textit{Embeddiogram}, offers a unique and powerful representation of an audio signal. The Embeddiogram is derived by applying a pre-trained neural network model to windowed segments of an audio signal, generating a sequence of embedding vectors. These embedding vectors collectively form a high-dimensional description of the audio signal's structure and content.

Below is a detailed explanation of computing the Embeddiogram from a given audio signal. This process comprises five key steps, including loading the audio data, slicing the audio data into windowed segments, processing each window using a pre-trained model to produce an embedding, collecting these embeddings, and finally normalizing the Embeddiogram.

\begin{enumerate}
\item \textbf{Load the audio data}: The audio data is loaded into memory as a one-dimensional array, denoted as $\text{input\_audio}$, of length $N$.

\item \textbf{Slice the audio data}: The audio data is segmented into overlapping windows. Each window contains $w$ samples, and consecutive windows are separated by $h$. This gives a total of $H$ windows, defined as:
\begin{equation}
H = 1 + \left\lfloor \frac{N - w}{h} \right\rfloor
\end{equation}
In the case where $\left( N - w \right) \mod h > 0$, we have $H += 1$ to account for the final, potentially smaller, window.

\item \textbf{Process each window}: Each window of audio data is processed independently, passed through the pre-trained neural network model and transformed into an embedding vector. Formally, for each window $w_i$ of audio data, we have:
\begin{equation}
\text{embedding}_i = \text{model}(w_i)
\end{equation}

\item \textbf{Collect the embeddings}: The embedding vectors are collected and stacked together. Each row represents a feature vector for a given time frame to form the \textit{Embeddiogram}, denoted as $\text{embeddiogram}$:
\begin{equation}
\text{embeddiogram} = \begin{bmatrix} \text{embedding}_1 \\ \text{embedding}_2 \\ \vdots \\ \text{embedding}_H \end{bmatrix}
\end{equation}

\item \textbf{Normalize the Embeddiogram}: The Embeddiogram is normalized to have a minimum value of 0 and a maximum value of 1. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

### 2023-05-14 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits the existence of high-level musical concepts invariant to specific attributes such as tonal quality, tempo, key, orchestration, and signal noise that evolve and unfolds through time. This idea parallels the nature of sheet music, which maintains its essence despite being interpreted through diverse performances using various instruments and styles.

Abstract high-level features like melody, harmony, rhythm, tempo, form, and expression can be discerned regardless of waveform or production style. This approach provides an objective, comprehensive insight into a composition's musical content and meaning, much like analyzing sheet music uncovers its structural and intentional foundation.

It's worth noting that while a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece of sheet music can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations. The distinct tonal quality of a waveform, heavily influenced by the specific timbre of the instrument or production technology, adds complexity to the transcription of its sonic properties into accurate sheet music notation.

Despite the differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose that figures \ref{fig:mahler} and \ref{fig:giant_steps} share a similarity. Both pieces carry the same musicological fingerprint, derived from the propulsion of their melodic and harmonic contours through non-diatonic major thirds.

This concept finds grounding in the theories of Kurt Koffka, a founder of the Gestalt school of psychology, who proposed a holistic approach to understanding complex forms, contrary to the structuralist practice of reducing mental processes into basic elements. His work, primarily embodied in the book, "Principles of Gestalt Psychology," established the foundation for Gestalt psychology and its applications to perception \cite{Koffka2013PrinciplesPsychology}.

>>>>

This thesis posits the existence of high-level musical concepts invariant to specific attributes such as tonal quality, tempo, key, orchestration, and signal noise that evolve and unfolds through time. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations. The distinct tonal quality of a waveform, heavily influenced by the specific timbre of the instrument or production technology, adds complexity to the transcription of its sonic properties into accurate sheet music notation.

Abstract high-level features can be discerned regardless of waveform or production style. This approach provides an objective, comprehensive insight into a composition's musical content and meaning, much like analyzing sheet music uncovers its intentional foundation.

For example, despite the differences in cultural contexts, stylistic elements, time signatures, key signatures, and sonic properties, we propose that figures \ref{fig:mahler} and \ref{fig:giant_steps} share a similarity. Both pieces carry the same musicological fingerprint so to say, derived from the propulsion of their melodic and harmonic contours through non-diatonic major thirds.

This concept finds grounding in the theories of Kurt Koffka, a founder of the Gestalt school of psychology, who proposed a holistic approach to understanding complex forms, contrary to the structuralist practice of reducing mental processes into essential elements \cite{Koffka2013PrinciplesPsychology}.

### 2023-05-14 introduction/Subsections/Objectives.tex
Updates from Overleaf

This study seeks to discover a generalization of the fundamental tonal structure in Western music by creating and examining an embedding space that can be applied to downstream Music Information Retrieval (MIR) tasks.

Neural networks have recently facilitated the development of low-dimensional musical embeddings that capture essential musical features, enabling the computation of content-based similarity between musical passages \cite{KimLEARNINGLOSS}\cite{Hung2022Feature-informedClassification}. An embedding function $f(x) \in \mathbb{R}^d$ maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, allowing the measurement of similarity between two musical passages using Euclidean distance, as denoted by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The success of such systems hinges on the level of musical abstraction and the embedding space's capacity to interpret music in a manner consistent with human perception. As a result, a suitable embedding space is vital for producing output that aligns with human expectations.

This research aims to thoroughly analyze musical works from various styles and genres, employing computational techniques from Music Information Retrieval (MIR) and a musicological perspective to contextualize the findings. Aural skills, encompassing perceptual and cognitive abilities, are necessary for understanding musical works through listening. The underlying musical structure consists of a systematic arrangement of elements that generate a coherent and expressive musical work. This study seeks to provide new insights into these components of musical composition and further our comprehension of their interrelationships.

We denote a single embedding for a music piece as $e(M)$.

For a given music piece $M$, we aim to learn a single embedding, $e(M)$, that captures both the underlying content (melody, harmony, critical, holistic elements) and production aspects. The embedding space should be structured to cluster similar content and production features together, effectively separating the two aspects. Formally, our objective is to learn the embedding function $e(M)$ such that:

\begin{itemize}
\item Similar content features are closer in the embedding space: $|e(M_1) - e(M_2)| < \epsilon_c$ for music pieces $M_1$ and $M_2$ with similar content.
\item Similar production features are closer in the embedding space: $|e(M_1) - e(M_3)| < \epsilon_p$ for music pieces $M_1$ and $M_3$ with similar production.
\item Distinct content and production features are further apart in the embedding space.
\end{itemize}

We propose learning a single embedding $e(M)$ that captures content and production features, with the embedding space structured to cluster similar aspects together, effectively distinguishing between the two.

Music is a complex and multi-dimensional phenomenon, and developing a model that emulates human aural skills is challenging. Nonetheless, recent advancements in deep learning offer promising avenues for creating a model that achieves satisfactory results in specific music-related tasks. This study will employ deep learning techniques to explore the correlation between aural skills and the underlying tonal structure of music.

Additionally, this research contends that MIR studies should transition from a solely technical-centered approach to a more balanced and comprehensive method that integrates other pertinent domains such as musicology and music theory. This shift will enable a better understanding of the interplay between technical and musical aspects of music, enhancing our comprehension of music signals and music itself.

>>>>

This study aims to explore the core tonal structure of Western music by constructing and analyzing an embedding space applicable to subsequent Music Information Retrieval (MIR) tasks.

By employing a blend of MIR computational techniques and a musicological approach, it intends to examine a wide spectrum of musical works across diverse styles and genres. Crucial to this is the understanding of aural skills – perceptual and cognitive abilities essential for interpreting musical works through listening.

The research focuses on the learning of a single embedding, $e(M)$, for a music piece $M$ which encapsulates the underlying high-level content such as melody contour, harmony inter-relationships, and other holistic elements. The structuring of the embedding space will prioritize the clustering of similar content and sonic quality features while maintaining a clear distinction between them.

The goal is to create an embedding $e(M)$ that captures high-level musical content and remains agnostic to production features.

### 2023-05-14 introduction/Subsections/Previous work.tex
Updates from Overleaf

\section{Previous work}

@@

\subsection{}

>>>>

\section{Related work}

@@

\subsection{XXXX}

Neural networks have recently facilitated the development of low-dimensional musical embeddings that capture essential musical features, enabling the computation of content-based similarity between musical passages \cite{Kim2021LearningLoss}\cite{Hung2022Feature-informedClassification}. An embedding function $f(x) \in \mathbb{R}^d$ maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, allowing the measurement of similarity between two musical passages using Euclidean distance, as denoted by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The success of such systems hinges on the level of musical abstraction and the embedding space's capacity to interpret music in a manner consistent with human perception. As a result, a suitable embedding space is vital for producing output that aligns with human expectations.
